---
title: "Emma's R Bootcamp"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome to bootcamp! 

This bootcamp was originally desgined for my MD/PhD cohort at UW-Madison. However, anyone can follow along! This curriculum is designed for individuals new to both R and coding in general. NO PRIOR CODING EXPERIENCE REQUIRED! **I have not created any materials presented in this bootcamp.** Rather, I have put together a roadmap of resources that I think will help students build the coding skills and statistical knowledge needed to analyze a large, "high-dimensional" dataset. 

Phase 1 consists entirely of CodeAcademy PRO's R course, in addition to practice sets designed by Jenny Bryan for UBC's STAT545 lecture. By the end of Phase 1 of this boot camp, you should be able to: 

* manipulate a dataset into "tidy" format
* generate a variety of different graphs from a "tidy" dataset
* analyze trends over different groupings of data
* calculate mean/median/mode/IQR and perform basic statistical tests (t.tests, generalized linear models)

After getting your feet wet with data cleaning/data viz, we will move on to more complex concepts featured in the R Software Carpentry workshops that will make your coding life easier! By the end of phase 2, you should be able to: 

* understand how to construct and interprete an if/else statement 
* push/pull something from github

Phase 3 of this bootcamp will focus on providing an overview of statistics concepts needed to understand most computational approaches, particularly those in the genomics field. We will focus particularly on two-group comparisons, linear modeling, multiple testing and supervised/unsupervised learning. Phase 3 will refer to slidedecks prepared for UBC's STAT540 course. By the end of phase 3, you should be able to: 

* understand what quality checks should be performed on a dataset before conducting any analyses
* understand the statistical underpinnings of two-group comparisons and linear models
* know why multiple testing is important in "big data" studies
* perform PCA analysis
* understand supervised and unsupervised clustering methods

##Syllabus 

###Phase 1

To complete this phase, you will need to sign up for CodeAcademy's PRO R course. After completing each lesson, do the practice sets in your Google collab document. If a lesson does not have any associated practice sets, continue on to the next lesson. If you have any questions, just make a comment in the Google Collab document and tag me. 

```{r, echo=FALSE}
phase1 <- read.csv("Phase1_syllabus.csv", stringsAsFactors = FALSE, check.names = FALSE)

knitr::kable(phase1)
```

###Phase 2

Make sure you have RStudio up and running on your own machine. Instructions for how to do this should have been provided in Lesson 1 of CodeAcademy, but if you didn't get a chance to set that up, there are instructions [here](https://happygitwithr.com/rstudio-git-github.html#rstudio-git-github).

```{r, echo=FALSE, warning=FALSE}
phase2 <- read.csv("Phase2_syllabus.csv", stringsAsFactors = FALSE, check.names = FALSE)
knitr::kable(phase2)
```

###Phase 3

This phase will focus on understanding conceptual topics foundational to genomics-based/statistical research. The slide decks below are referenced from UBC's STAT540 course. They are used with permission from Sara Mostafavi (course director). 

```{r, echo=FALSE, warnings=FALSE}
phase3 <- read.csv("Phase3_syllabus.csv", stringsAsFactors = FALSE, check.names = FALSE)
knitr::kable(phase3)
```





